Optimal (Multiway) Spatial Joins

### 2024-09-19 msj.tex
yufei

Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$. Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$ and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle. We define
\myeqn{
    B_\bm{t} &=& \bigcap_{i=1}^{\red{k}} r_i; \label{eqn:B_t}
}
namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, define:
\myitems{
    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.

@@

Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ. To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time. We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$. Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$. We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.

@@

The rest of the section will focus on how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(n \log n + k \cdot \out)$ time. An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ within the same time complexity.

@@

Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, we maintain the set $R_\ell$ of {\em active} rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$. Specifically, when $\ell$ hits the bottom edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert $[x_1, x_2]$ into $\T$ and mark $r$ as active, which can be done in $O(\log n)$ time. The rectangle $r$ will be referred to as the {\em host} of $[x_1, x_2]$ and is stored together with $[x_1, x_2]$ in $\T$.

>>>>

Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$. Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$ and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle. We define
\myeqn{
    B_\bm{t} &=& \bigcap_{i=1}^{\red{k}} r_i; \label{eqn:B_t}
}
namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, if $B_\bm{t}$ is not empty, define:
\myitems{
    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.

@@

Our objective is to prove that H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ (note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ). To ensure the soundness of our notation system, let us formulate the ``1-SJ'' as the trivial problem where the input is a set $R$ of $n$ rectangles, and the goal is simply to enumerate each rectangle of $R$; the problem can obviously be ``solved'' in $O(n)$ time. We assume the existence of an algorithm $\A$ that can settle $\kappa$-SJ for all $\kappa \in [1, k-1]$. Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$. We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$ for any $\kappa \ge 1$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.

@@

The rest of the section will focus on how to prepare the sets $\dcross_H(\bm{t})$ of all $\bm{t} \in \J(R_1', ..., R_{k-2}')$ in $O(k n \log n + k \cdot \out)$ time. An analogous method can be used to compute the sets $\dcross_V(\bm{t})$ of all $\bm{t}$ within the same time complexity.

@@

Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, we maintain the set $R_\ell$ of {\em active} rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$. Specifically, when $\ell$ hits the top edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert $[x_1, x_2]$ into $\T$ and mark $r$ as active, which can be done in $O(\log n)$ time. The rectangle $r$ will be referred to as the {\em host} of $[x_1, x_2]$ and is stored together with $[x_1, x_2]$ in $\T$.
